### 2019-02-19 paper_3.tex

%%
%% This is file `example/ch_intro.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% install/buptgraduatethesis.dtx  (with options: `ch-intro')
%% 
%% This file is a part of the example of BUPTGraduateThesis.
%%

\chapter{基于动态阈值的pairwise注意力机制模型}

\section{引言}

\section{DPAM算法设计}

\subsection{Pairwise attention model}

\subsection{Dynamic threshold predictor}

\subsection{模型学习与预测}

\section{实验及分析}

\subsection{数据集介绍}

\subsection{模型评估}

\subsection{实验结果}

\subsubsection{Perforamnce of two sub-models}

\subsubsection{Comparison aginst two sub-models}

\subsubsection{Comaprison against baselines}

\subsection{分析与讨论}

% \chapterbib