### 2019-02-21 paper_4.tex

%%
%% This is file `example/ch_intro.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% install/buptgraduatethesis.dtx  (with options: `ch-intro')
%% 
%% This file is a part of the example of BUPTGraduateThesis.
%%

\chapter{基于递归注意力机制的模型}

\section{引言}

\section{DPAM算法设计}

\subsection{编码层}

\subsection{自注意力层}

\subsection{递归层}

\subsection{输出层}

\subsection{模型学习与预测}

\section{实验设置}

\subsection{数据集介绍}

\subsection{模型评估}

\section{实验结果}

\subsection{Comparison against Baselines}

\section{分析与讨论}

\subsection{The impact of recurrent layer}

\subsection{Ablation test}

% \chapterbib