### 2019-02-22 paper_6.tex

%%
%% This is file `example/ch_intro.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% install/buptgraduatethesis.dtx  (with options: `ch-intro')
%% 
%% This file is a part of the example of BUPTGraduateThesis.
%%

\chapter{基于BERT的多任务学习模型}

\section{引言}
\label{sec:ran_intro}

\section{多任务学习框架}